Mathematical Aspects of the Capital Market

### 2017-08-08 phd/Chapterintroduction.tex
Corrected the introduction chapter of the thesis.

\section{Regular Variation}
The concept of regular variation is defined by the following scaling
property: if a function $f(\cdot)$ satisfies
\[
\lim_{x \to \infty} {f(c x) \over f(x)} = c^\alpha
\quad
\forall c > 0
\]
then we say $f(x)$ is regularly varying with index $\alpha$.
% Distribution functions, say $F(\cdot)$, whose survival function
% $\bar F(x) = 1 - F(x)$ satisfies the above scaling property, are
% observed in a variety of phenomena, 
When expanded to multiple dimensions, the scaling property of regular
variation is better described in terms of vague convergence to a
spectral measure $\mu_\alpha$: if a random vector $V$ satisfies
\[
{
  \P(V/|V| \in \cdot, |V| > c x)
  \over
  \P(|V| > x)
},
\overset{v}{\to} c^{-\alpha} \mu_\alpha(\cdot)
\quad
x \to \infty, \forall c > 0
\]
then we say the $V$ is regularly varying with index $\alpha$. Here
$\mu_\alpha(\cdot)$ is a probability measure on the unit sphere
\cite{buraczewski:damek:mikosch:2016}. Clearly, if $V$
is regularly varying with index $\alpha$, then each component
of it and each linear combination of its components are regularly
varying with the same index $\alpha$. This follows from Feller
\cite{feller}, p. 278. Cf. also Jessen and Mikosch
\cite{JessenMikosch2006}, lemma 3.1, and Embrechts et. al.
\cite{embrechts:klueppelberg:mikosch:1997}, lemma 1.3.1.

Clearly, estimating the tail index $\alpha$ is particularly important
for understanding the behaviour of a heavy-tailed series. A standard
method proposed for this purpose is due to Hill \cite{hill1975simple}:
\[
\hat \alpha_H = \left[
  {1 \over k} \sum_{i=1}^k \log \left(
  X_{(i)} \over X_{(k+1)}
  \right)
  \right]^{-1}
\]
where $X_1, X_2, \dots$ is the series whose tail index is the subject
of estimation, and $X_{(i)}$ is the $i$-th upper order statistics of
this series. Several authors have contributed to showing the
weak consistency of the estimator $\hat \alpha_H$,
i.e. if $k \to \infty, k/n \to 0$ as $n \to \infty$, then
$\hat \alpha_H \overset{P}{\to} \alpha$.

Mason \cite{mason:1982} first proved that the estimator
was consistent when $X_1, X_2, \dots$ were iid; later Rootz\'en,
Leadbetter and de Haan \cite{rootzen:leadbetter:dehaan1992} and Hsing
\cite{hsing:1991} proved its consistency when $X_1, X_2, \dots$ were
weakly dependent; and Resnick and St{\u a}ric{\u a}
\cite{resnick:starica:1995, resnick:starica:1997} proved its
consistency when $X_t$ was a linear process.

@@

Now consider $N$ return series $X_{i,t}$, $i=1,2,\dots, p$. Suppose
each of these series is a linear combination of $K$ factors $Y_{i,t}$,
$i=1,2,\dots,K$, the $i$-th of which is regularly varying with index
$\alpha_i$. Then by the summation property, each and every $X_i$ is
regularly varying with index $\min_{1 \leq i \leq K} \alpha_i$.
Since in practice a factor $Y_{i,t}$ is found as a linear combination
of $X_{i,t}$, the observed time series, following an eigenvector of
the sample covariance matrix of $X_{i,t}$, it is important to
understand the eigensystem of this matrix. This topic indeed
constitutes chapters \ref{ch:bernoulli} and \ref{ch:extremes} of this
thesis.

>>>>

\section{Regular Variation}
The concept of regular variation is defined by the following scaling
property: if a function $f$ on $(0, \infty)$ satisfies
\[
\lim_{x \to \infty} {f(c x) \over f(x)} = c^\alpha
\quad
\forall c > 0
\]
then we say $f$ is regularly varying with index $\alpha$.
$f$ can be written in the form $f(x) = {L(x) \over x^\alpha}$, where
$L(x)$ is a slowly varying function, i.e.
$\lim_{x \to \infty} {L(c x) \over L(x)} = 1$, $\forall c > 0$.
We call a random variable $X$ regularly varying with index
$\alpha \geq 0$ if it satisfies the tail balance condition
\[
\P(X > x) \sim p_+ {L(x) \over x^\alpha},\quad
\P(X < -x) \sim p_- {L(x) \over x^\alpha}, \quad
\text{for some } p_+, p_- > 0,\; p_+ + p_- = 1
\]

% Distribution functions, say $F(\cdot)$, whose survival function
% $\bar F(x) = 1 - F(x)$ satisfies the above scaling property, are
% observed in a variety of phenomena, 
When expanded to multiple dimensions, the scaling property of regular
variation is better described in terms of vague convergence to a
spectral measure $\mu_\alpha$: if a random vector $V$ satisfies
\[
{
  \P(V/|V| \in \cdot, |V| > c x)
  \over
  \P(|V| > x)
}
\overset{v}{\to} c^{-\alpha} \mu_\alpha(\cdot)
\quad
x \to \infty, \forall c > 0,
\]
then we say $V$ is regularly varying with index $\alpha$. Here
$\mu_\alpha$ is a probability measure on the unit sphere
\cite{buraczewski:damek:mikosch:2016}. Clearly, if $V$
is regularly varying with index $\alpha$, then each component
and each linear combination of its components are regularly
varying with the same index $\alpha$. This follows from Feller
\cite{feller}, p. 278. Cf. also Jessen and Mikosch
\cite{JessenMikosch2006}, lemma 3.1, and Embrechts et. al.
\cite{embrechts:klueppelberg:mikosch:1997}, lemma 1.3.1.

Clearly, estimating the tail index $\alpha$ of a sequence
$X_1, X_2, \dots$ of regularly varying variables is particularly
important for understanding the behaviour of a heavy-tailed series. A
standard method proposed for this purpose is due to Hill
\cite{hill1975simple}:
\[
\hat \alpha_H = \left[
  {1 \over k} \sum_{i=1}^k \log \left(
  X_{(i)} \over X_{(k+1)}
  \right)
  \right]^{-1}
\]
where $X_1, X_2, \dots, X_n$ is a sample whose tail index is the
subject of estimation, and $X_{(i)}$ is the $i$-th upper order
statistic of the sample. Several authors have contributed to showing
the weak consistency and asymptotic normality of the estimator
$\hat \alpha_H$, under the assumptions $k \to \infty, k/n \to 0$
as $n \to \infty$.

%% Mason \cite{mason:1982} first proved that the estimator
%% was consistent when $X_1, X_2, \dots$ were iid; later Rootz\'en,
%% Leadbetter and de Haan \cite{rootzen:leadbetter:dehaan1992} and Hsing
%% \cite{hsing:1991} proved its consistency when $X_1, X_2, \dots$ were
%% weakly dependent; and Resnick and St{\u a}ric{\u a}
%% \cite{resnick:starica:1995, resnick:starica:1997} proved its
%% consistency when $X_t$ was a linear process.

@@

Now consider $p$ return series $X_{i,t}$, $i=1,2,\dots, p$. Suppose
each of these series is a linear combination of $K$ factors $Y_{j,t}$,
$j=1,2,\dots,K$, the $j$-th of which is regularly varying with index
$\alpha_i$. Then by the summation property, each and every $X_i$ is
regularly varying with index $\min_{1 \leq i \leq K} \alpha_i$.
Since in practice a factor $Y_{i,t}$ is found as a linear combination
of $X_{i,t}$, the observed time series, following an eigenvector of
the sample covariance matrix of $X_{i,t}$, it is important to
understand the eigensystem of this matrix. This topic indeed
constitutes chapters \ref{ch:bernoulli} and \ref{ch:extremes} of this
thesis.
